---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
Loading Libraries
```{r LoadLibraries}
      library(knitr)
      library(sqldf)
```

Setting Global Echo to TRUE
```{r setoptions, echo=TRUE}
      opts_chunk$set(echo=TRUE)
```

## Loading and preprocessing the data
```{r LoadData}
      activity <- read.csv("activity.csv")
      activity$steps <- as.numeric(activity$steps)
      activity$date <- as.Date(activity$date, "%Y-%m-%d")
      activity$interval <- as.numeric(activity$interval)
      #head(activity)
      #str(activity)
      #summary(activity)
```

## What is mean total number of steps taken per day?

####Histogram 
```{r StepsDayHist}
      df <- sqldf(
            "select *
            from activity
            where steps is not null")

      df2 <- sqldf(
            "select date, sum(steps) AS tot_steps
            from df
            group by date")

      with(df2, plot(date, tot_steps, ylab = "total steps", type = "h"))
```

####Mean and Median of Total Steps per Day
```{r StepsDayCalc}
      df3 <- sqldf(
            "select date, avg(steps) AS mean_steps, median(steps) AS median_steps
            from df
            group by date")
      df3
```

## What is the average daily activity pattern?

####Times Series 
```{r AvgDailyActivity}
      df4 <- sqldf(
            "select interval, avg(steps) AS mean_steps
            from df
            group by interval")

      with(df4, plot(interval, mean_steps, ylab = "mean steps", type = "l"))
```

####Intereval with the Maximum Number of Steps
```{r AvgDailyMax}
      df5 <- sqldf(
            "select interval, mean_steps AS steps
            from df4
            where mean_steps =
                  (select max(mean_steps)
                  from df4)")

      df5
```

## Imputing missing values

####Total Missing Values
```{r MissTotal}
      df6 <- sqldf(
            "select *
                  from activity
                  where steps is null")

      nrow(df6)
```

####Imputing Missing Values with Interval Mean
```{r MissComp}
activity2 <- sqldf(
      "select 
            case
                  when a.steps is null then mean_steps
                  else a.steps 
            end AS steps, 
            a.date,
            a.interval
      from activity a
            left outer join df4 d on a.interval= d.interval")
```

####Histogram with Imputed Data
```{r StepsDayHist2}
      df7 <- sqldf(
            "select date, sum(steps) AS tot_steps
            from activity2
            group by date")

      with(df7, plot(date, tot_steps, type = "h"))
```

####Mean and Median of Total Steps per Day with Imputed Data
```{r StepsDayCalc2}
      df8 <- sqldf(
            "select date, avg(steps) AS mean_steps, median(steps) AS median_steps
            from activity2
            group by date")
      df8
```

## Are there differences in activity patterns between weekdays and weekends?

####Creating Day Type Variable
```{r DayType}
      day_type <- weekdays(activity2$date)

      df9 <- cbind(activity2, day_type)

      df10 <- sqldf(
            "select steps,
            date,
            interval,
            case
                  when day_type in ('Saturday', 'Sunday') then 'weekend'
                  when day_type is null then null
                  else 'weekday'
            end as day_type
            from df9")
```

####Panel Plot
```{r DayTypePlot}
df11_wday <- sqldf(
      "select interval, avg(steps) as mean_steps
      from df10
      where day_type = 'weekday'
      group by interval")

df11_wend <- sqldf(
      "select interval, avg(steps) as mean_steps
      from df10
      where day_type = 'weekend'
      group by interval")

par(mfcol = c(2, 1))

with(df11_wend, plot(interval, mean_steps, type = "l", main="weekend", xlab = "Interval", ylab="Avg Number of Steps"))

with(df11_wday, plot(interval, mean_steps, type = "l", main="weekday", xlab="Interval", ylab="Avg Number of Steps"))

par(mfcol = c(1, 1))
```
